---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#R> ",
  fig.path = "man/figures/README-"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(didnp)
# library(ggpubr)
# data(mpdta)
# devtools::load_all("~/Dropbox/did")
# library(BMisc)
```
# A model-free differences-in-differences framework

<!-- badges: start -->

```{r echo=FALSE, results='asis', message=FALSE}
cat(
  badger::badge_cran_download("npsf", "grand-total", "blue"),
  badger::badge_cran_download("npsf", "last-month", "yellow"),
  badger::badge_cran_release("npsf", "green"),
  badger::badge_devel("OlegBadunenko/didnp", "red"),
  badger::badge_cran_checks("npsf"),
  # badger::badge_codecov("OlegBadunenko/did"),
  badger::badge_last_commit("OlegBadunenko/didnp")
)
```

<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

The **didnp** package contains tools for computing average treatment effect parameters in a Difference-in-Differences setup allowing for

* this
* that

The main parameters are **that**.  These are 


## Getting Started

The **didnp** package implements the framework developed in

* 


## Installation

You can install **didnp** from CRAN with:

```{r eval=FALSE}
install.packages("didnp")
```

or get the latest version from github with:

```{r gh-installation, eval = FALSE}
if ( !require("devtools") ) install.packages("devtools"); library(devtools)
devtools::install_github("OlegBadunenko/didnp")
```

## An example

The following is a simplified example ...,  which comes from [that](https://elsevier.com/).

A subset of the data is available in the package and can be loaded by
```{r eval=FALSE}
  library(didnp)
  data(qqqqqqqq)
```

The dataset contains XXX observations

* **LHS**  lhs variable

To estimate verage treatment effects, use the **didnpreg** function

```{r eval=FALSE}
tym1 <- didnpreg(
  form1,
  data = dat2,
  subset = mysmpl,
  bwmethod = "opt",
  boot.num = 399,
  TTb = TRUE,
  print.level = 2,
  cores = 32)
```

**didnpreg** returns a class **didnp** object.  This object contains estimates of the average treatment effects and their standard errors.  To see these, we can call the **summary** function

```{r eval=FALSE}
summary(tym1)
```

This provides estimates of 

It is often also convenient to plot the group-time average treatment effects.  This can be done using the **didnpplot** command:

```{r echo=FALSE}
library(ggplot2)
```

```{R eval=FALSE}
didnpplot(tym1, ylim = c(-.25,.1))
```

The ... in the plot are 

## Additional Resources

That 
