---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#R> ",
  fig.path = "man/figures/README-"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(didnp)
# library(ggpubr)
# data(mpdta)
# devtools::load_all("~/Dropbox/did")
# library(BMisc)
```

# A model-free differences-in-differences framework

<!-- badges: start -->

```{r echo=FALSE, results='asis', message=FALSE}
cat(
  badger::badge_cran_download("npsf", "grand-total", "blue"),
  badger::badge_cran_download("npsf", "last-month", "yellow"),
  badger::badge_cran_release("npsf", "green"),
  badger::badge_devel("OlegBadunenko/didnp", "red"),
  badger::badge_cran_checks("npsf"),
  # badger::badge_codecov("OlegBadunenko/did"),
  badger::badge_last_commit("OlegBadunenko/didnp")
)
```

<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

The **didnp** package contains tools for computing average treatment effect parameters in a Difference-in-Differences setup without specifying a model.  


## The framework

The **didnp** package implements the framework developed in

Daniel J. Henderson and Stefan Sperlich (2023), "A Complete Framework for Model-Free Difference-in-Differences Estimation", Foundations and Trends® in Econometrics: Vol. 12: No. 3, pp 232-323. http://dx.doi.org/10.1561/0800000046


## Installation

The package **didnp** can be install from CRAN by typing:

```{r eval=FALSE}
install.packages("didnp", dependencies = TRUE)
```

or get the latest version from github by typing:

```{r gh-installation, eval = FALSE}
if ( !require("devtools") ) install.packages("devtools"); library(devtools)
devtools::install_github("OlegBadunenko/didnp")
```

## Illustration

## Data

> The subset of the data from [IPUMS](https://ipums.org) is used to illustrate the functionality of the package. The data are available in the package and can be loaded by

```{r data, eval = TRUE}
  library(didnp)
  library(tidyverse)
  data(Unempl, package = "didnp")
  head(Unempl)
```

The description of the dataset can be found by typing

```{r datahelp, eval = FALSE}
  ?Unempl
  table(Unempl$age)
```

The variable 'ELIGIBLE' is actually treated. Only those who were eligible before 2012 can be treated. In 2013 and beyond, eligible individuals are treated. Conversely, in 2011 and earlier, they are eligible to be treated starting from 2013.

```{r treated, eval = TRUE}
table(Unempl$YEAR, Unempl$ELIGIBLE)
```

For convenience, a data said with a smaller number of variables is created

```{r smallLetters, eval = TRUE}
d2 <- data.frame(
  y = as.numeric(Unempl$FT),
  year = Unempl$YEAR,
  unemp = Unempl$UNEMP,
  sex = factor(Unempl$SEX),
  age = ordered(Unempl$AGE),
  yrimmig = ordered(Unempl$YRIMMIG),
  educ = ordered(Unempl$EDUC),
  treatment_period = as.numeric(Unempl$AFTER),
  treated = as.numeric(Unempl$ELIGIBLE)
)
```

There is a gap in the 'year'

```{r gapYear, eval = TRUE}
table(d2$year)
```

Year 2012 is missing. This is when the treatment occurred. A variable 't' is generated, where

 | Year | t | 
 | :---  |  -: | 
 | 2008 | -3 | 
 | 2009 | -2 | 
 | 2010 | -1 | 
 | 2011 | 0 | 
 | 2013 | 1 | 
 | 2014 | 2 | 
 | 2015 | 3 | 
 | 2016 | 4 | 

An artificial variable 'period' that does not have a gap is generated, where

 | Year | period | 
 | :---  |  -: | 
 | 2008 | 1 | 
 | 2009 | 2 | 
 | 2010 | 3 | 
 | 2011 | 4 | 
 | 2013 | 5 | 
 | 2014 | 6 | 
 | 2015 | 7 | 
 | 2016 | 8 | 

To make the same more homogeneous and smaller for illustrative purposes, the subsample of the data with the following conditions is created:

1. yrimmig >= 1982 & yrimmig <= 1994
2. educ >= 6 & educ <= 11


```{r period, eval=TRUE}
year_min <- min(Unempl$YEAR)

d0 <- d2 %>%
  filter(yrimmig >= 1982 & yrimmig <= 1994) |>
  filter(educ >= 6 & educ <= 11) |>
  mutate(
    yrimmig = droplevels(yrimmig),
    educ = droplevels(educ),
    age = droplevels(age)
    ) |>
    mutate(period = dplyr::if_else(year < 2012,
      year - year_min + 1,
      year - year_min),
      t = period - min(period[treatment_period == 1]-1))
table(d0$period)
```

> Here is a basic description of the data

Treated over time
```{r Treated_t, eval=TRUE}
table(d0$year, d0$treated)
table(d0$t, d0$treated)
```

Recoded variable 't'

```{r tYear, eval=TRUE}
table(d0$t, d0$year)
```

Treated by the year of immigration
```{r TreatedYrimmig, eval=TRUE}
table(d0$yrimmig, d0$treated)
```

Treated by education
```{r TreatedEduc, eval=TRUE}
table(d0$educ, d0$treated)
```

Treated by age
```{r TreatedAge, eval=TRUE}
table(d0$treated, d0$age)
```

Although this can be done on the fly, the subsample can be prepared beforehand:

```{r sub, eval=TRUE}
# get the subsample
d0 <- d0 %>%
  mutate(smpl = year >= 2010 & year <= 2014)
table(d0$smpl)
```

> Define the formula that we will use:

```{r formula, eval = TRUE}
form1 <- y ~ age + educ + sex + unemp | period | treated | treatment_period
```

To obtain standard errors and perform testing in this *illustration*,

> we will use a few number of bootstrap replicaitons here, but we advise to set `boot.num = 399` or larger in an application. 


```{r boot, eval = TRUE}
B <- 9
```

## Testing

To test if there is a violation of the bias stability condition use command **didnptest**

```{r model1test, eval = TRUE}
tym1test <- didnpbsctest(
  form1,
  data = d0,
  subset = smpl,
  boot.num = B,
  print.level = 2,
  cores = 10)
```

*Interpretation*: We do find evidence against the null hypothesis that the bias stability condition holds. The desired *p*-value should be much larger than the 0.1 level.

## Estimation

To estimate the average treatment effects, we use the **didnpreg** function.  The **didnpreg** function allows using matrices. The manual explains how to use matrix syntax (type `?didnpreg`). 

To speed up the estimation

> on computers with multiple cores, use multiplrocessing by setting option `cores`. 

Suppress output by setting `print.level = 0`. The default value is 1.

```{r model1a, eval=TRUE}
# suppress output
tym1a <- didnpreg(
  form1,
  data = d0,
  subset = smpl,
  bwmethod = "opt",
  boot.num = B,
  TTx = "TTa",
  print.level = 2,
  digits = 8,
  cores = 10)
```

**didnpreg** returns a class **didnp** object.  This object contains estimates of the average treatment effects and their standard errors.  To see these, we can call the **summary** function.

```{r model1asumm, eval=TRUE}
# Print the summary of estimation
summary(tym1a)
rm(tym1a)
```

Estimating $TT_b$ will take longer. The bandwidths are cross-validated.

```{r model1b, eval=TRUE}
# Show output as the estimation goes
tym1b <- didnpreg(
  form1,
  data = d0,
  subset = smpl,
  bwmethod = "CV",
  boot.num = B,
  TTx = "TTb",
  print.level = 2,
  digits = 4,
  cores = 10)
```

### Understanding samples

> To work with the results for the plot, **didnpreg** produces binary variables indicating different samples.

The estimation sample is captured by the *esample* value, which is equal to 1 if this observation was used in the estimation and 0 otherwise:
```{r esample, eval=TRUE}
table(
  d0$year[tym1b$esample],
  d0$treated[tym1b$esample]
)
```


The value *sample1* indicates **all treated** (before and after treatment) observations in the estimation sample *esample*. To subset the treated observations in the original dataset, double subsetting is required.
```{r esamplesample1, eval=TRUE}
table(
  d0$year[tym1b$esample][tym1b$sample1],
  d0$treated[tym1b$esample][tym1b$sample1]
)
```

The value *sample11* indicates the **treated** observations in the post-treatment period in the estimation sample *esample*. To subset the treated observations in the original dataset, double subsetting is required.
```{r esamplesample11, eval=TRUE}
table(
  d0$year[tym1b$esample][tym1b$sample11],
  d0$treated[tym1b$esample][tym1b$sample11]
)
# or the same in case TTx = "TTa" was used by the 'didnpreg' command
table(
  d0$year[tym1b$esample][tym1b$sample1][tym1b$TTa.positions.in.TTb],
  d0$treated[tym1b$esample][tym1b$sample1][tym1b$TTa.positions.in.TTb]
)
```

The values of the individual TEs can be put to the dataset using double subsetting for TTb
```{r esamplesample1data, eval=TRUE}
d0$TTbi[tym1b$esample][tym1b$sample1] <- tym1b$TTb.i
d0 %>%
  group_by(year) %>%
  summarize(mean = mean(TTbi, na.rm = TRUE))
```

or TTa

```{r esamplesample11data, eval=TRUE}
d0$TTai[tym1b$esample][tym1b$sample11] <- tym1b$TTa.i
d0 %>%
  group_by(year) %>%
  summarize(mean = mean(TTai, na.rm = TRUE))
```

> Just a remainder that TTa calculates the ATET in the post-treatment period, while TTb calculated the ATET in both prior and post-treatment periods.

## Plotting Heterogenous Treatment Effects

To plot the heterogenous treatment effects, use the **didnpplot** command. 

> The heterogenous treatment effects can be plotted **by** either continuous or categorical variable. They can also also be plotted **by** either continuous or categorical variable **over** another categorical variable.

### 'by': factor

#### education

The heterogenous treatment effects are plotted for each level of education, since the 'education' is categorical variable.

```{r tym1b_gr_educ, eval=TRUE}
tym1b_gr_educ <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$educ[tym1b$esample][tym1b$sample1],
  xlab = "Education",
  ylab = "ATET"
)
# A
tym1b_gr_educ$data.a
tym1b_gr_educ$plot.a
# B
tym1b_gr_educ$data.b
tym1b_gr_educ$plot.b
# ggsave(paste0("atet_ci_education.pdf"), width = 15, height = 10, units = c("cm"))
```

Here objects `data.a` and `data.b` contain data that is used to produce `plot.a` and `plot.b`. The graphs are `ggplot` objects and can be amended further.

#### time

The heterogenous treatment effects over time show that the effect is significant in the second year after the treatment.

> Note that the graph shows the **90%** confidence interval bygiving the option `level = 90`.


```{r tym1b_gr_time, eval=TRUE}
tym1b_gr_time <- didnpplot(
  obj = tym1b,
  level = 90,
  by = factor(d0$t)[tym1b$esample][tym1b$sample1],
  xlab = "Time",
  ylab = "ATET"
)
# A
tym1b_gr_time$data.a
tym1b_gr_time$plot.a
# B
tym1b_gr_time$data.b
tym1b_gr_time$plot.b
# ggsave(paste0("atet_ci_time.pdf"), width = 15, height = 10, units = c("cm"))
```

#### sex

Another example is the graph with the heterogenous treatment effects by sex:

```{r tym1b_gr_sex, eval=TRUE}
tym1b_gr_sex <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$sex[tym1b$esample][tym1b$sample1],
  xlab = "Sex",
  ylab = "ATET",
  by.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_sex$data.a
tym1b_gr_sex$plot.a
# B
tym1b_gr_sex$data.b
tym1b_gr_sex$plot.b
# ggsave(paste0("atet_ci_sex.pdf"), width = 15, height = 10, units = c("cm"))
```

### 'by' continuous: unemp

When the 'by' variable is continuous (the **didnpplot** command will recognize if `by` is a continuous variable) the **didnpplot** offers three ways of plotting the heterogeneous treatment effects.

#### Scale of the continuous 'by' is its range

If `by.continuous.scale` is not specified, **didnpplot** will use each unique value in the 'by' variable to produce the plot, which can be pretty rugged.

```{r tym1b_gr_unemp_each_value, eval=TRUE}
tym1b_gr_unemp_each_value <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  xlab = "Unemployment",
  ylab = "ATET",
  add.zero.line = FALSE
)
# A
head(tym1b_gr_unemp_each_value$data.a, 10)
tym1b_gr_unemp_each_value$plot.a
# B
head(tym1b_gr_unemp_each_value$data.b, 10)
tym1b_gr_unemp_each_value$plot.b
# ggsave(paste0("atet_ci_unemp_numeric.pdf"), width = 15, height = 10, units = c("cm"))
```

Ameding ggplot object is easy. For example adding a 0 horizontal line is

```{r gr_age2, eval=TRUE}
tym1b_gr_unemp_each_value$plot.a +
  geom_hline(yintercept = 0)
```

> Anternatively, one can use the `data.a` and `data.b` objects to plot from scratch.

#### Scale of the continuous 'by' intervals

Setting `by.continuous.scale` to a single number, instructs **didnpplot** to split the range of the continuous ‘by’ variable into the number of intervals specified by `by.continuous.scale` and plot ATET by intervals.

```{r tym1b_gr_unemp_n_intervals, eval=TRUE}
tym1b_gr_unemp_n_intervals <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  by.continuous.scale = 10,
  xlab = "Unemployment",
  ylab = "ATET",
  by.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_unemp_n_intervals$data.a
tym1b_gr_unemp_n_intervals$plot.a
# B
tym1b_gr_unemp_n_intervals$data.b
tym1b_gr_unemp_n_intervals$plot.b
# ggsave(paste0("atet_ci_unemp_intervals.pdf"), width = 15, height = 10, units = c("cm"))
```

#### Scale of the continuous 'by' a vector

Setting `by.continuous.scale` to a single number, instructs **didnpplot** to split the range of the continuous ‘by’ variable into intervals defined by the specified vector and plot ATET by intervals.

```{r tym1b_gr_unemp_vector_breaks, eval=TRUE}
tym1b_gr_unemp_vector_breaks <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  by.continuous.scale = c(2, 3, 4, 5, 7, 9, 12),
  xlab = "Unemployment",
  ylab = "ATET",
  by.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_unemp_vector_breaks$data.a
tym1b_gr_unemp_vector_breaks$plot.a
# B
tym1b_gr_unemp_vector_breaks$data.b
tym1b_gr_unemp_vector_breaks$plot.b
# ggsave(paste0("atet_ci_unemp_breaks.pdf"), width = 15, height = 10, units = c("cm"))
```

### 'by': factor + 'over

> The heterogeneous treatment effects can be plotted by **either** continuous or categorical variable over a **categorical** variable.

#### 'by': sex; 'over' is time

This is an example with treatment effects by sex over time.

```{r tym1b_gr_sex_time, eval=TRUE}
tym1b_gr_sex_time <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$sex[tym1b$esample][tym1b$sample1],
  over = as.factor( d0$t[tym1b$esample][tym1b$sample1]),
  # over = d0$educ[tym1b$esample][tym1b$sample1],
  xlab = "Sex",
  ylab = "ATET",
  by.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_sex_time$data.a
tym1b_gr_sex_time$plot.a
# B
tym1b_gr_sex_time$data.b
tym1b_gr_sex_time$plot.b
# ggsave(paste0("atet_ci_sex_time.pdf"), width = 15, height = 10, units = c("cm"))
```

#### 'by': time; 'over' is sex

This is an example with treatment effects by time over sex, reversing the order of the previous graph. 

```{r tym1b_gr_time_sex, eval=TRUE}
tym1b_gr_time_sex <- didnpplot(
  obj = tym1b,
  level = 95,
  over = d0$sex[tym1b$esample][tym1b$sample1],
  by = as.factor( d0$t[tym1b$esample][tym1b$sample1]),
  # over = d0$educ[tym1b$esample][tym1b$sample1],
  xlab = "Time",
  ylab = "ATET",
  over.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_time_sex$data.a
tym1b_gr_time_sex$plot.a
# B
tym1b_gr_time_sex$data.b
tym1b_gr_time_sex$plot.b
# ggsave(paste0("atet_ci_time_sex.pdf"), width = 15, height = 10, units = c("cm"))
```

> Alternatively use the data from the object `tym1b_gr_time_sex`

to produce another type of graph:

```{r gr_age_race2, eval=TRUE}
crit.value <- 2
pd <- position_dodge(0.1) # move them .05 to the left and right
d1 <- tym1b_gr_time_sex$data.b
d1$Sex <- d1$over
ggplot(d1, aes(x = by, y = atet, color = Sex, group = Sex)) +
  geom_errorbar(aes(ymin = atet - crit.value*atet.sd, ymax = atet + crit.value*atet.sd), color = "black", width = .1, position = pd) +
  geom_line(position = pd) +
  geom_point(position = pd, size = 3, shape = 21, fill = "white") +
  xlab("Time") +
  ylab("ATET") +
  theme_bw() +
  theme(legend.position = "right", text = element_text(size = 17))
```

### 'by' continuous: unemp + 'over'

The following examples use continuous 'by' variable to plot the heterogeneous treatment facts over sex.

#### Scale of the continuous 'by' is its range

`by.continuous.scale` is not specified or `NULL`.

```{r tym1b_gr_unemp_each_value_sex, eval=TRUE}
tym1b_gr_unemp_each_value_sex <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  over = d0$sex[tym1b$esample][tym1b$sample1],
  xlab = "Unemployment",
  ylab = "ATET",
  over.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
head(tym1b_gr_unemp_each_value_sex$data.a)
tym1b_gr_unemp_each_value_sex$plot.a
# B
head(tym1b_gr_unemp_each_value_sex$data.b)
tym1b_gr_unemp_each_value_sex$plot.b
# ggsave(paste0("atet_ci_unemp_sex_numeric.pdf"), width = 15, height = 10, units = c("cm"))
```

#### Scale of the continuous 'by' intervals

`by.continuous.scale` is a scalar.

```{r tym1b_gr_unemp_intervals_sex, eval=TRUE}
tym1b_gr_unemp_intervals_sex <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  by.continuous.scale = 15,
  over = d0$sex[tym1b$esample][tym1b$sample1],
  xlab = "Unemployment",
  ylab = "ATET",
  over.labels.values = data.frame(c("1","2"), c("Male", "Female"))
)
# A
tym1b_gr_unemp_intervals_sex$data.a
tym1b_gr_unemp_intervals_sex$plot.a
# B
tym1b_gr_unemp_intervals_sex$data.b
tym1b_gr_unemp_intervals_sex$plot.b
# ggsave(paste0("atet_ci_unemp_sex_intervals.pdf"), width = 15, height = 10, units = c("cm"))
```

#### Scale of the continuous 'by' a vector

`by.continuous.scale` is a vector. Note even if the user provides an interval wider than the range of the continuous 'by' variable **didnpplot** will take care of that and define the plausible range for plotting. In this example, `print.level = 2` option is used to show the working of the  **didnpplot** command.


```{r tym1b_gr_unemp_breaks_sex, eval=TRUE}
tym1b_gr_unemp_breaks_sex <- didnpplot(
  obj = tym1b,
  level = 95,
  by = d0$unemp[tym1b$esample][tym1b$sample1],
  by.continuous.scale = seq(0, 15, 1), #c(2, 5, 7, 9, 12),
  over = d0$sex[tym1b$esample][tym1b$sample1],
  xlab = "Unemployment",
  ylab = "ATET",
  over.labels.values = data.frame(c("1","2"), c("Male", "Female")),
  print.level = 2
)
# A
tym1b_gr_unemp_breaks_sex$data.a
tym1b_gr_unemp_breaks_sex$plot.a
# B
tym1b_gr_unemp_breaks_sex$data.b
tym1b_gr_unemp_breaks_sex$plot.b
# ggsave(paste0("atet_ci_unemp_sex_breaks.pdf"), width = 15, height = 10, units = c("cm"))
```

## Additional Resources

To be added



