---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#R> ",
  fig.path = "man/figures/README-"
)
```

```{r, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
library(didnp)
# library(ggpubr)
# data(mpdta)
# devtools::load_all("~/Dropbox/did")
# library(BMisc)
```
# A model-free differences-in-differences framework

<!-- badges: start -->

```{r echo=FALSE, results='asis', message=FALSE}
cat(
  badger::badge_cran_download("npsf", "grand-total", "blue"),
  badger::badge_cran_download("npsf", "last-month", "yellow"),
  badger::badge_cran_release("npsf", "green"),
  badger::badge_devel("OlegBadunenko/didnp", "red"),
  badger::badge_cran_checks("npsf"),
  # badger::badge_codecov("OlegBadunenko/did"),
  badger::badge_last_commit("OlegBadunenko/didnp")
)
```

<!-- badges: end -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

The **didnp** package contains tools for computing average treatment effect parameters in a Difference-in-Differences setup without specifying a model.  


## Getting Started

The **didnp** package implements the framework developed in

* 


## Installation

You can install **didnp** from CRAN with:

```{r eval=FALSE}
install.packages("didnp", dependencies = TRUE)
```

or get the latest version from github with:

```{r gh-installation, eval = FALSE}
if ( !require("devtools") ) install.packages("devtools"); library(devtools)
devtools::install_github("OlegBadunenko/didnp")
```

## An example

> The following is a simplified example ...,  which comes from [that](https://elsevier.com/).

Data set is shared by [Kuka et al, 2020](https://doi.org/10.1257/pol.20180352). Here we showcase the functionality of the package by using a subset, which is available in the package and can be loaded by
```{r data, eval = TRUE}
  library(didnp)
  library(ggplot2)
  data(DACAsub, package = "didnp")
  head(DACAsub)
```

The description of the dataset can be found by typing

```{r datahelp, eval = FALSE}
  ?DACAsub
```

Although this can be done on the fly, the subsample can be prepared beforehand:

```{r sub, eval=TRUE}
# get the subsample
DACAsub$mysmpl <- mysmpl <-
  DACAsub$a1922==1 & !is.na(DACAsub$a1922) &
  DACAsub$htus==1 & !is.na(DACAsub$htus)

# generate 'treatment_period'
DACAsub$treatment_period <- ifelse( DACAsub[,"year"] > 2011, 1, 0)
```

To estimate the average treatment effects, we use the **didnpreg** function. Define the formula that we will use:

```{r formula, eval = TRUE}
form1 <- inschool ~ fem + race + var.bpl + state + age + yrimmig + 
  ageimmig | inschool | year | elig | treatment_period | perwt
```

The **didnpreg** function allows using matrices. The manual explains how to use matrix syntax (type `?didnpreg`). To obtain standard errors,

> we will use a few number of bootstrap replicaitons here, but we advise to set `boot.num = 399` or more. 

To speed up the estimation

> on computers with multiple cores, use multiplrocessing by setting option `cores`. 

Suppress output by setting `print.level = 0`. The default value is 1.

```{r model1a, eval=TRUE}
# suppress output
tym1a <- didnpreg(
  form1,
  data = DACAsub,
  subset = mysmpl,
  bwmethod = "opt",
  boot.num = 7,
  TTb = FALSE,
  print.level = 0,
  cores = 8)
```

**didnpreg** returns a class **didnp** object.  This object contains estimates of the average treatment effects and their standard errors.  To see these, we can call the **summary** function.

```{r model1asumm, eval=TRUE}
# Print the summary of estimation
summary(tym1a)
```

Estimating $TT_b$ will take longer. The bandwidths is cross-validated.


```{r model1b, eval=TRUE}
# Show output as the estimation goes
tym1b <- didnpreg(
  form1,
  data = DACAsub,
  subset = mysmpl,
  bwmethod = "CV",
  boot.num = 99,
  TTb = TRUE,
  print.level = 1,
  cores = 8)
```

To plot the heterogenous treatment effects, use the **didnpplothte** command. Define three variables by and over which the treatment effects will be plotted:

```{r by_over, eval=TRUE}
DACAsub[tym1b$esample, "race"] -> race
DACAsub[tym1b$esample, "fem"]  -> sex
as.numeric(DACAsub[tym1b$esample, "age"]) -> age
```

Here `age` is quasi-continuous. It will have only 4 values (in fact fewer than the Race variable), but we use it to showcase the functionality.

First, use one categorical `by` variable:

```{r gr_race, eval=TRUE}
tym1b_gr_race <- didnpplothte(
  obj = tym1b,
  level = 95,
  by = race[tym1b$sample1],
  xlab = "Race",
  ylab = "ATET",
  by.labels.values = data.frame(
    old = c(1,2,3,4,5),
    new = c("Hispanic", "White", "Black", "Asian", "Other")
  ))

tym1b_gr_race$data.a
tym1b_gr_race$data.b

tym1b_gr_race$plot.a
tym1b_gr_race$plot.b
```

Here objects `data.a` and `data.b` contain data that is used to produce `plot.a` and `plot.b`. The graphs are `ggplot` objects and can be amended further.

Another example is the graph with treatment effects by sex. Note the sample object `tym1b` is used:

```{r gr_sex, eval=TRUE}
tym1b_gr_sex <- didnpplothte(
  obj = tym1b,
  level = 95,
  by = sex[tym1b$sample1],
  xlab = "Sex",
  ylab = "ATET",
  by.labels.values = data.frame(c(1,0), c("Female", "Male"))
)

tym1b_gr_sex$data.a
tym1b_gr_sex$data.b

tym1b_gr_sex$plot.a
tym1b_gr_sex$plot.b
```

The **didnpplothte** command will recognize if `by` is continuous variable, split it into given number of intervals `n.intervals` and plot the treatment effects by the split variable.

```{r gr_age, eval=TRUE}
tym1b_gr_age <- didnpplothte(
  obj = tym1b,
  level = 95,
  by = age[tym1b$sample1],
  n.intervals = 10,
  xlab = "Age"
)

tym1b_gr_age$data.a
tym1b_gr_age$data.b

tym1b_gr_age$plot.a
tym1b_gr_age$plot.b
```

Ameding ggplot object is easy. For example adding a 0 horizontal line is

```{r gr_age2, eval=TRUE}
tym1b_gr_age$plot.a +
  geom_hline(yintercept = 0)
```

> Anternatively, one can use the `data.a` and `data.b` objects to plot from scratch.

The treatment effects can be visualized by and over. Variable `by` can be both categorical and continuous, while `over` must be categorical. For example, to plot treatment effects by age over race, specify the `over` option:

```{r gr_age_race, eval=TRUE}
tym1b_gr_age_race <- didnpplothte(
  obj = tym1b,
  level = 90,
  by = age[tym1b$sample1],
  n.intervals = 7,
  over = race[tym1b$sample1],
  xlab = "Age",
  ylab = "ATET",
  point_size = 2,
  over.labels.values = data.frame(
    old = c(1,2,3,4,5),
    new = c("Hispanic", "White", "Black", "Asian", "Other")
  ),
  text_size = 15)

tym1b_gr_age_race$data.a
tym1b_gr_age_race$plot.a

tym1b_gr_age_race$data.b
tym1b_gr_age_race$plot.b
```

Note that the graph shows the **90%** confidence interval. 

> Alternatively use the data from the object `tym1b_gr_age_race`

to produce another type of graph:

```{r gr_age_race2, eval=TRUE}
crit.value <- 2
pd <- position_dodge(0.1) # move them .05 to the left and right
d1 <- tym1b_gr_age_race$data.b
d1$Race <- d1$over
ggplot(d1, aes(x = by, y = atet, color = Race, group = Race)) +
  geom_errorbar(aes(ymin = atet - crit.value*atet.sd, ymax = atet + crit.value*atet.sd), color = "black", width = .1, position = pd) +
  geom_line(position = pd) +
  geom_point(position = pd, size = 3, shape = 21, fill = "white") +
  xlab("Age") +
  ylab("ATET") +
  theme_bw() +
  theme(legend.position = "right", text = element_text(size = 17))
```

The next plot shows treatment effects by age over sex:

```{r gr_age_sex, eval=TRUE}
tym1b_gr_age_sex <- didnpplothte(
  obj = tym1b,
  level = 90,
  by = age[tym1b$sample1],
  n.intervals = 7,
  over = sex[tym1b$sample1],
  xlab = "Age",
  ylab = "ATET",
  point_size = 2,
  over.labels.values = data.frame(c(1,0), c("Female", "Male")),
  text_size = 15)

tym1b_gr_age_sex$data.a
tym1b_gr_age_sex$plot.a

tym1b_gr_age_sex$data.b
tym1b_gr_age_sex$plot.b
```

Finally, both `by` and `over` are both categorical:

```{r gr_sex_race, eval=TRUE}
tym1b_gr_sex_race <- didnpplothte(
  obj = tym1b,
  level = 95,
  by = sex[tym1b$sample1],
  over = race[tym1b$sample1],
  xlab = "Sex",
  ylab = "ATET",
  point_size = 3,
  by.labels.values = data.frame(c(1,0), c("Female", "Male")),
  over.labels.values = data.frame(
    old = c(1,2,3,4,5),
    new = c("Hispanic", "White", "Black", "Asian", "Other")
  ),
  text_size = 17)

tym1b_gr_sex_race$data.a
tym1b_gr_sex_race$plot.a

tym1b_gr_sex_race$data.b
tym1b_gr_sex_race$plot.b
```

## Additional Resources

That 
